---
title: "Calculating BIST Net Income From OECD Data"
author: "Ian Salter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
library(pracma)
library(EnvStats)
source("../R/distributions.R")
```

The [OECD lists](https://www.oecd.org/social/income-distribution-database.htm) Gini and Mean Income data for  before and after the application of tax and benefits. From this data, assuming a given probability distribution, it is possible to derive a simple model of a net income function that shows the net income for very possible gross income value. This approximation is named a Basic Income Simple Tax (BIST) net income function. 

Here we consider two possible probability distributions as possible models of income distribution: The Log Normal Distribution and the Pareto Distribution.

The [Log Normal Distribution](https://en.wikipedia.org/wiki/Log-normal_distribution) is defined in terms of two parameters $\mu$, the mean, and $\sigma$, the standard deviation of the variables natural logarithm (sdlog). The following plot shows the Probability Density Function for the logormal distribution with $\mu = 0$ and various values of $\sigma$: 

```{r echo = FALSE}
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
p +
  stat_function(fun = dlnorm, args = list(sdlog = 0.25), colour = "red") +
  geom_text(aes(x= 2.2, y = 1.5), label = "sdlog = 0.25", size = 3, colour = "red") +
  stat_function(fun = dlnorm, args = list(sdlog = 0.5), colour = "green") +
  geom_text(aes(x= 2.5, y = 0.5), label = "sdlog = 0.5", size = 3, colour = "green") +
  stat_function(fun = dlnorm, args = list(sdlog = 1), colour = "blue") +
  geom_text(aes(x= 2.8, y = 0.25), label = " sdlog = 1", size = 3, colour = "blue") +
  xlim(0, 5) +
  ylim(0, 2) +
  ylab("PDF")
```

The [Pareto Distribution](https://en.wikipedia.org/wiki/Pareto_distribution) is defined in terms of two parameters $x_{m}$, the scale (or location) and $\alpha$, the shape, The following plot shows the Probability Density Function for the pareto distribution for $x_{m} = 1$ and various values of $\alpha$:

```{r echo = FALSE}
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
p +
  stat_function(fun = dpareto, args = list(location = 1, shape = 3), colour = "red") +
  geom_text(aes(x= 2.2, y = 1.5), label = "shape = 3", size = 3, colour = "red") +
  stat_function(fun = dpareto, args = list(location = 1, shape = 2), colour = "green") +
  geom_text(aes(x= 2.5, y = 0.5), label = "shape = 2", size = 3, colour = "green") +
  stat_function(fun = dpareto, args = list(location = 1, shape = 1), colour = "blue") +
  geom_text(aes(x= 3.5, y = 0.25), label = " shape = 1", size = 3, colour = "blue") +
  xlim(0, 5) +
  ylim(0, 3) +
  ylab("PDF")
```  


Given values of $gini$ and mean $\mu$ we can calculate the parameters for both lognormal and pareto distributions For the lognormal distribution:

$\mu = \mu$

$\sigma = 2erf^{-1}(gini)$

For the pareto distribution (assuming gini is not 1):

$x_{m} = \frac{(\alpha - 1)\mu}{\alpha}$

$\alpha = \frac{1 + gini}{2gini}$














For 2017 we have the following OECD data for the UK:

```{r echo = FALSE}
total_earnings = 20051
capital_income = 1223
self_emp_income = 3156

gini_net = 0.357
mean_net = 22879
gini_gross = 0.506
mean_gross = total_earnings + capital_income + self_emp_income
```

* Gini (disposable income, post taxes and transfers): `r format(gini_net)` 
* Mean disposable income (current prices): `r format(mean_net)`   
* Gini (market income before taxes and transfers) : `r format(gini_gross)`  
* Mean Income Before Tax: `r format(mean_gross)` calculated by summing  
  + Total Earnings (current prices): `r format(total_earnings)`
  + Capital Income (current prices): `r format(capital_income)`
  + Income from self-employment and goods produced for own consumption (current prices): 
  `r format(self_emp_income)`




  
Figure 1 shows log normal probability distributions for the UK in 2017 where the OECD data:

```{r echo = FALSE}
lnorm_sdlog(gini_net)
probability_distribution <- function(x){
  dlnorm(x)
}
modified_distribution <- function(x){
  dlnorm(x, sdlog = lnorm_sdlog(gini_net))
}

p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
p +
  stat_function(fun = probability_distribution) +
  stat_function(fun = modified_distribution,
                mapping = aes(linetype = "dotted")) +
                xlim(0, 5) +
                ylim(0, 1)
```

```{r echo = FALSE}
net_income_distribution <- function(x){
  dlnorm(x, meanlog = mean_net, sdlog = lnorm_sdlog(gini_net))
}
gross_income_distribution <- function(x){
  dlnorm(x, meanlog = mean_gross, sdlog = lnorm_sdlog(gini_gross))
}
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
p +
  stat_function(fun = net_income_distribution) +
  stat_function(fun = gross_income_distribution) +
  xlim(0, 1000000) +
  ylim(0, 0.000000000000000000001)
```

Stuff below is from template

---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
